### 2020-08-17 article.tex
first init

\documentclass{article}

\usepackage[round]{natbib}
\usepackage{hyperref}
\usepackage{dirtree}
\usepackage{graphicx}
\graphicspath{{img/}}

\title{ARGoS-Blockchain interface}
\author{Volker Strobel \and Alex Pacheco \and Marco Dorigo}

\begin{document}
\maketitle

\section{Introduction}

Several recent works have addressed the combination of blockchain
technology and swarm robotics. This combinatin proves useful for
security, consensus finding, robot economies, robot-as-a-service, and
potentially many more.

However, running these experiments requires a suitable framework. Such
a framework proves useful both for researchers who do not have access
to a suitable robot platform, such as the Pi-puck robot. Additionally,
a suitable platform is important before running time- and
cost-consuming experiments with real robots.

The ARGoS robot simulator \citep{PinTriOGr-etal2012:si} is the
state-of-the-art research platform to conduct simulations in swarm
robotics.

In general, in swarm robotics research, each robot is a blockchain
node, and contributes to the maintenance of the system by exchanges
blockchain information. To link ARGoS and blockchain software, we
developed the ARGoS-Blockchain interface that provides access to the
blockchain nodes for the robots (Figure~\ref{fig:interface}). The
interface is intended to facilitate research in blockchain-based robot
swarms by allowing to call blockchain functions in
ARGoS. Additionally, Docker makes it easy to install and run the
interface on different platforms.

\section{Software availability}

%The Blockchain module is available at
%\url{https://github.com/Pold87/ARGoS-Blockchain-interface} and the
%ARGoS module is available at
%\url{https://github.com/Pold87/ARGoS-Blockchain-interface}.

%\dirtree{%
%.1 geth/.
%.1 img/.
%.1 local_scripts/.
%.1 .gitignore.
%.1 README.
%.1 docker-compose.yml.
%}

% TODO: this is a good figure, also include it
% \begin{figure}
%   \centering
%     \includegraphics[width=\textwidth]{Figure5}
%   \caption{For each robot, the ARGoS-Blockchain interface establishes
%     a connection to the Ethereum blockchain via a Docker container and
%     shell scripts that provide templates for executing Ethereum (geth)
%     functions.}
%   \label{fig:interface}  
% \end{figure}

\section{Overview}

The implementation of the custom Ethereum network is based on
Capgemini \textsc{aie}'s Ethereum
Docker\footnote{\url{https://github.com/Capgemini-AIE/ethereum-docker},
  accessed on November 6, 2019}. Docker containers
\citep{Mer2014:linux} contain all the necessary dependencies to run
specific applications and are more lightweight than a virtual
machine. In our setup, for each robot, the Ethereum implementation
\emph{geth} is executed in a separate Docker container. The simulated
robots maintain a \emph{custom} Ethereum network, i.e., a network that
is shared among the simulated robots and independent of Ethereum's
main network. Different containers can communicate with each other via
channels.

In order to execute an Ethereum function (e.g., create a new smart
contract) from ARGoS, a robot uses its C\texttt{++} controller to
attach to the Docker container. The Docker containers provide shell
scripts\footnote{The interface uses shell scripts, since, during
  development, it became evident that they are executed much faster
  than other Ethereum APIs.} with customizable templates (e.g., one of
the templates compiles the smart contract, uses the binary code to
send a blockchain transactions, and waits until the contract is
mined). Via Ethereum's \textsc{ipc} (interprocess communications)
interface, the shell scripts execute the Ethereum functions.

We use an auxiliary `bootstrap' node for publishing the smart contract
to the blockchain at the beginning of each run of the simulations
(Figure~\ref{fig:initialization}). The bootstrap node then mines the
smart contract and sends the contract address and the \textsc{abi}
(application binary interface; the \textsc{abi} specifies which
functions a smart contract provides and how to call them) to the
controllers of the robots. As soon as this is done, the bootstrap node
is removed from the network. The bootstrap node is not necessarily
required and the smart contract could also be created by a
robot. However, we used an auxiliary node to make sure (i) that the
smart contract is available at the start of the actual experimental
run and (ii) that robots have the same initial conditions in all
experiments.

The experiments were conducted on a computer cluster. To simulate the
limited hardware of real robots, one core with 2.0~GHz and 1.8~GB of
memory was assigned to each Docker container\footnote{This is a
  reasonable choice as a robot's computer could easily have such
  characteristics. It is also a convenient choice because on a
  computer with 2.0~GHz and 1.8~GB of RAM, Ethereum works
  ``out-of-the-box,'' without any modifications; therefore, any
  interested user can obtain the most recent release of Ethereum from
  the official depository and use it with our publicly available
  ARGoS-Blockchain interface.}. The communication channels between the
Docker containers were only established when robots were within a
50~cm communication range in order to simulate the local communication
capabilities of real robots.

The description in this report is intended to allow everyone to run
blockchain-based robot experiments in simulation. There is no need for
specific hardware but powerful CPU and RAM are advantageous for fluent
experiments.

The ARGoS-Blockchain interface is composed of two modules.

\subsubsection{Module~1: Blockchain part}
Module~1 allows for creating an Ethereum network with several nodes,
where each node is located in a separate Docker container. For the
ARGoS-Blockchain interface, the interaction with the Ethereum nodes is
done via C\texttt{++}, using the code in the following repository:
\url{https://github.com/Pold87/robot-swarms-need-blockchain}

This repository contains one module of the ARGoS-Blockchain interface
that is described in the article ``Blockchain Technology Secures Robot
Swarms: A Comparison of Consensus Protocols and Their Resilience to
Byzantine Robots by Strobel, V., Castello Ferrer, E., and Dorigo, M.''

For debugging purposes or for creating your own private Ethereum
network, you can also use this module without ARGoS.

\begin{figure}[t]
  \centering
  \includegraphics{interface}
  \caption[Overview of the ARGoS-Blockchain interface]{This scheme
    gives an overview of the workflow of the ARGoS-Blockchain
    interface}
  \label{fig:interface}
\end{figure}

\subsubsection{Module~2: ARGoS part}

\section{Setup}

The very first time you run this code, it is required to create the
Docker image for the Ethereum nodes and initialize Docker Swarm as
follows:

\begin{verbatim}
cd geth/
docker build -t mygeth .
docker swarm init
\end{verbatim}

Additionally, you have to set the variable \verb|DOCKERFOLDER| in the
file \verb|global_config.sh| to the full path where this repository is
located on your computer, for example:

\begin{verbatim}
/Users/vstrobel/Documents/ARGoS-Blockchain-interface/
\end{verbatim}

In order to be able to mine, you need to create the DAG datasets as
follows (the creates files require approximately 2~GB disk space and
the execution of the script can take take several minutes):

\begin{verbatim}
cd local_scripts/
bash create_dag.sh
\end{verbatim}

\section{Run}
Usually, the network is created when a swarm robotics experiment is
started, using one of the start scripts in
\url{https://github.com/Pold87/robot-swarms-need-blockchain}.

However, you can also start the Ethereum network without ARGoS, using
the following command:

\begin{verbatim}
bash local_scripts/start_network.sh <number of nodes>
\end{verbatim}

That is, \verb|bash local_scripts/start_network.sh 5|, would create a
private Ethereum network with 5 nodes.

\subsubsection*{Acknowledgements}

Volker Strobel and Marco Dorigo acknowledge support from the Belgian
F.R.S.-FNRS, of which they are a Research Fellow and a Research
Director respectively. Alexandre Pacheco acknowledges support from the
Faculty of Applied Sciences of the Universit\'{e} Libre de Bruxelles.

%\bibliographystyle{splncs04}
\bibliography{library}

\end{document}

### 2020-08-25 article.tex
upd

@@
 Several recent works have addressed the combination of blockchain
-technology and swarm robotics. This combinatin proves useful for
-security, consensus finding, robot economies, robot-as-a-service, and
-potentially many more.
-
-However, running these experiments requires a suitable framework. Such
-a framework proves useful both for researchers who do not have access
-to a suitable robot platform, such as the Pi-puck robot. Additionally,
-a suitable platform is important before running time- and
-cost-consuming experiments with real robots.
-
-The ARGoS robot simulator \citep{PinTriOGr-etal2012:si} is the
-state-of-the-art research platform to conduct simulations in swarm
-robotics.
+technology and swarm robotics
+\cite{StrCasDor2018:aamas,StrDor2018:ants,StrFerDor2017:techreport-013}. This
+combination proves useful for security, consensus finding, robot
+economies, robot-as-a-service, and potentially many other applications
+in robot swarms. However, conducting experiments involving robot
+swarms and blockchain technology requires a suitable testbed. A widely
+used framework consists of the Pi-puck robot platform running the
+Ethereum blockchain network. However, not every researcher has access
+to physical robots, obtaining and maintaining them be both
+time-intense and cost-intense, not to mention the lengthy execution of
+experiments. Therefore, a simulation testbed is needed to quickly try
+and test different controllers and scenarios. The ARGoS robot
+simulator \citep{PinTriOGr-etal2012:si} is the state-of-the-art
+research platform to conduct simulations in swarm robotics. Such a
+framework proves useful both for researchers who do not have access to
+a suitable robot platform, such as the Pi-puck robot. Additionally, a
+suitable platform is important before running time- and cost-consuming
+experiments with real robots.
 
@@
 swarms by allowing to call blockchain functions in
-ARGoS. Additionally, Docker makes it easy to install and run the
+ARGoS. Additionally, using Docker makes it easy to install and run the
 interface on different platforms.
@@
 
-%The Blockchain module is available at
-%\url{https://github.com/Pold87/ARGoS-Blockchain-interface} and the
-%ARGoS module is available at
-%\url{https://github.com/Pold87/ARGoS-Blockchain-interface}.
+The open-source software is available at GitHub: the blockchain module
+is available at
+\footnote{\url{https://github.com/Pold87/ARGoS-Blockchain-interface}}
+and the ARGoS module is available at
+\footnote{\url{https://github.com/Pold87/robot-swarms-need-blockchain}}.
 
-%\dirtree{%
-%.1 geth/.
-%.1 img/.
-%.1 local_scripts/.
-%.1 .gitignore.
-%.1 README.
-%.1 docker-compose.yml.
-%}
-
-% TODO: this is a good figure, also include it
-% \begin{figure}
-%   \centering
-%     \includegraphics[width=\textwidth]{Figure5}
-%   \caption{For each robot, the ARGoS-Blockchain interface establishes
-%     a connection to the Ethereum blockchain via a Docker container and
-%     shell scripts that provide templates for executing Ethereum (geth)
-%     functions.}
-%   \label{fig:interface}  
-% \end{figure}
+\begin{figure}[t]%\centering
+\dirtree{%
+.1 geth/.
+.1 img/.
+.1 local\_scripts/.
+.1 .gitignore.
+.1 README.
+.1 docker-compose.yml.
+}
+\label{fig:folderstructure}  
+\caption{The folder structure of the package.}
+\end{figure}
+\begin{figure}
+  \centering
+    \includegraphics[width=\textwidth]{overview}
+  \caption{For each robot, the ARGoS-Blockchain interface establishes
+    a connection to the Ethereum blockchain via a Docker container and
+    shell scripts that provide templates for executing Ethereum (geth)
+    functions.}
+  \label{fig:overview}  
+\end{figure}
 
@@
 
+\subsection{Requirements}
+\begin{itemize}
+\item Linux
+\item Docker
+\end{itemize}
+
+\subsection{Installation: Blockchain module}
 The very first time you run this code, it is required to create the
@@
 \begin{verbatim}
-cd geth/
-docker build -t mygeth .
-docker swarm init
+git clone https://github.com/Pold87/ARGoS-Blockchain-interface
+cd ARGoS-Blockchain-interface/ && cd geth/
+sudo docker build -t mygeth .
+sudo docker swarm init
 \end{verbatim}
 
-Additionally, you have to set the variable \verb|DOCKERFOLDER| in the
+\subsection{Installation: ARGoS module}
+
+\begin{verbatim}
+git clone https://github.com/Pold87/robot-swarms-need-blockchain
+cd robot-swarms-need-blockchain/
+mkdir cmake/
+cd cmake/
+cmake ..
+make
+\end{verbatim}
+
+Additionally, you have to set the variable \verb|DOCKERBASE| in the
 file \verb|global_config.sh| to the full path where this repository is
@@
 \begin{verbatim}
-/Users/vstrobel/Documents/ARGoS-Blockchain-interface/
+DOCKERBASE='/home/vstrobel/Documents/docker-geth-network/'
 \end{verbatim}
@@
 
-%\bibliographystyle{splncs04}
+\bibliographystyle{plain}
 \bibliography{library}

### 2020-09-02 article.tex
upd

@@
 \maketitle
+
+\begin{abstract}
+
+  This technical report documents the installation and use of the
+  ARGoS-Blockchain interface. This interface allows for conducting
+  experiments in the robot swarm simulator ARGoS and the smart
+  contract blockchain framework Ethereum. The blockchain part uses the
+  container software Docker in order to facilitate the installation
+  and the parallelization of the blockchain nodes. The use of
+  blockchain-based smart contracts provides a framework for managing
+  security issues in robot swarms. The interface is a convenient way
+  to prototype blockchain-based control software in robot
+  swarms. Byzantine robots, that is, robots that behave in an
+  unintended way can disrupt the swarm behavior. In this technical
+  report, we detail the motivation, setup, and usage of the
+  ARGoS-blockchain interface.
+
+\end{abstract}
 
@@
 
+The following setup steps describe the installation using Linux as
+operating system.
+
 \subsection{Requirements}
 \begin{itemize}
-\item Linux
+\item Sufficient disk space (at least 16 GB)
 \item Docker 
+\item ARGoS\footnote{\url{https://github.com/ilpincy/argos3}} and the
+  ARGoS e-puck robot
+  plugin\footnote{\url{https://github.com/demiurge-project/argos3-epuck}}
+  (carefully read the instructions on the respective respositories)
 \end{itemize}
 
-\subsection{Installation: Blockchain module}
-The very first time you run this code, it is required to create the
-Docker image for the Ethereum nodes and initialize Docker Swarm as
-follows:
+\subsection{Installation --- Module~1: Blockchain module}
+
+1. First, obtain the source code for the blockchain module:
+
+\begin{verbatim}
+$ git clone https://github.com/Pold87/ARGoS-Blockchain-interface
+\end{verbatim}
+
+2. Then, you need to create the Docker image for the Ethereum nodes and
+initialize Docker Swarm as follows. By default, the execution of
+Docker commands requires root user privileges. To facilitate the
+execution of Docker commands, it is possible to run Docker as non-root
+user (i.e., without \texttt{sudo}) as follows (see
+\url{https://docs.docker.com/engine/install/linux-postinstall/} for
+more detailed instructions):
+
+\begin{verbatim}
+$ sudo groupadd docker
+$ sudo usermod -aG docker $USER
+\end{verbatim}
+
+4. Then, create the Ethereum image.
+
+\begin{verbatim}
+$ cd ARGoS-Blockchain-interface/ && cd geth/
+$ docker build -t mygeth .
+$ docker swarm init
+\end{verbatim}
+
+5. Set the ARGoS folder in start_network.sh
+
+TODO: Run script with NUMROBOTS=0
+
+TODO: Change from python to python3 in Dockerfile
+
+TODO: Download geth and put it into myowngeth/
+
+TODO:
+
+\subsection{Installation: ARGoS module}
+
+1. Obtain the source code for the ARGoS module.
 
@@
 git clone https://github.com/Pold87/ARGoS-Blockchain-interface
-cd ARGoS-Blockchain-interface/ && cd geth/
-sudo docker build -t mygeth .
-sudo docker swarm init
 \end{verbatim}
 
-\subsection{Installation: ARGoS module}
-
+2. Compile the ARGoS code.
 \begin{verbatim}
-git clone https://github.com/Pold87/robot-swarms-need-blockchain
 cd robot-swarms-need-blockchain/
-mkdir cmake/
-cd cmake/
+mkdir build/
+cd build/
 cmake ..
@@
 
-Additionally, you have to set the variable \verb|DOCKERBASE| in the
-file \verb|global_config.sh| to the full path where this repository is
-located on your computer, for example:
+3. Set the variable \verb|DOCKERBASE| in the file
+\verb|global_config.sh| to the full path where the Blockchain module
+(Module 1) repository is located on your computer, for example:
 
@@
 
-In order to be able to mine, you need to create the DAG datasets as
+4. In order to be able to mine, you need to create the DAG datasets as
 follows (the creates files require approximately 2~GB disk space and
@@
 \section{Run}
+
 Usually, the network is created when a swarm robotics experiment is
@@
 
+\section{Debugging}
+
+First, check that the Docker container for the bootstrap node and the
+individual containers for the robots are running:
+
+\begin{verbatim}
+
+\end{verbatim}
+
+% Mention --no-trunc and how to access logs
+
+\section{Usage}
+
+The functions of the interface can be found in the file
+\texttt{/interface/generic_interface.cpp}.
+
+\begin{itemize}
+\item \texttt{unlockAccount}: Unlock the coinbase account.
+\item \texttt{startMining}: Start the mining process using one thread.
+\item \texttt{stopMining}: Stop the mining process.
+\item \texttt{addPeer}: Add a peer based on its enode.
+\item \texttt{removePeer}: Add a peer based on its enode.
+\item \texttt{scInterface}: 
+\end{itemize}
+
 \subsubsection*{Acknowledgements}

### 2020-09-03 article.tex
upd

@@
 
+\usepackage{listings}
+\usepackage{enumitem}
+\lstset{language=bash} 
 \usepackage[round]{natbib}
@@
 ARGoS-Blockchain interface, the interaction with the Ethereum nodes is
-done via C\texttt{++}, using the code in the following repository:
+done via C\texttt{++}, using the code in the following repository:\\
 \url{https://github.com/Pold87/robot-swarms-need-blockchain}
@@
   \centering
-  \includegraphics{interface}
+  \includegraphics[width=\textwidth]{interface}
   \caption[Overview of the ARGoS-Blockchain interface]{This scheme
@@
 
-\subsection{Installation --- Module~1: Blockchain module}
+\subsection{Obtain Source Code}
 
-1. First, obtain the source code for the blockchain module:
+\begin{enumerate}[leftmargin=*]
+\item Create a base folder for storing the two modules of the
+  ARGoS-Blockchain interface.
+\begin{verbatim}
+$ mkdir interface/
+$ cd interface/
+$ pwd
+/home/vstrobel/software/interface
+\end{verbatim}
 
+\item Obtain the source code for the blockchain module.
 \begin{verbatim}
@@
 
-2. Then, you need to create the Docker image for the Ethereum nodes and
-initialize Docker Swarm as follows. By default, the execution of
-Docker commands requires root user privileges. To facilitate the
-execution of Docker commands, it is possible to run Docker as non-root
-user (i.e., without \texttt{sudo}) as follows (see
+\item Obtain the source code for the ARGoS module.
+\begin{verbatim}
+git clone https://github.com/Pold87/robot-swarms-need-blockchain"
+\end{verbatim}
+\end{enumerate}
+
+\subsection{Installation --- Module~1: Blockchain module}
+
+\begin{enumerate}[leftmargin=*]
+\item By default, the execution of Docker commands requires root user
+  privileges. To facilitate the execution of Docker commands, it is
+  possible to run Docker as non-root user (i.e., without
+  \texttt{sudo}) as follows (see
   \url{https://docs.docker.com/engine/install/linux-postinstall/} for
-more detailed instructions):
-
+  more detailed instructions).
 \begin{verbatim}
@@
 \end{verbatim}
-
-4. Then, create the Ethereum image.
+ % New item
+\item Obtain the Ethereum source code and store it in the folder
+\verb|myowngeth/|. Please note that Ethereum is compiled locally
+here. While there are Docker images for Ethereum, this allows for more
+customization (e.g., if you want to change the size of the DAG file,
+as explained in Section~\ref{sec:limited-hardware}).
 
@@
 $ cd ARGoS-Blockchain-interface/ && cd geth/
+$ git clone https://github.com/ethereum/go-ethereum.git myowngeth/
+\end{verbatim}
+% New item
+\item Create the Ethereum image and initialize the Docker swarm.
+
+\begin{verbatim}
 $ docker build -t mygeth .
@@
 \end{verbatim}
-
-5. Set the ARGoS folder in start_network.sh
-
-TODO: Run script with NUMROBOTS=0
-
-TODO: Change from python to python3 in Dockerfile
-
-TODO: Download geth and put it into myowngeth/
-
-TODO:
-
-\subsection{Installation: ARGoS module}
-
-1. Obtain the source code for the ARGoS module.
+% New item
+\item Set the ARGoS folder in the file \verb|start_network.sh|.
 
 \begin{verbatim}
-git clone https://github.com/Pold87/ARGoS-Blockchain-interface
+$ nano ../local_scripts/start_network
+ARGOSFOLDER="/home/vstrobel/software/interface/robot-swarms-need-blockchain"
+\end{verbatim}
+% TODO: Run script with NUMROBOTS=0
+\end{enumerate}
+\subsection{Installation --- Module~2: ARGoS module}
+
+\item Compile the ARGoS code.
+\begin{verbatim}
+$ cd /home/vstrobel/software/interface/robot-swarms-need-blockchain/
+$ mkdir build/
+$ cd build/
+$ cmake ..
+$ make
 \end{verbatim}
 
-2. Compile the ARGoS code.
-\begin{verbatim}
-cd robot-swarms-need-blockchain/
-mkdir build/
-cd build/
-cmake ..
-make
-\end{verbatim}
-
-3. Set the variable \verb|DOCKERBASE| in the file
+\item Set the variable \verb|DOCKERBASE| in the file
 \verb|global_config.sh| to the full path where the Blockchain module
@@
 \begin{verbatim}
+$ cd ..
+$ nano global_config.sh
 DOCKERBASE='/home/vstrobel/Documents/docker-geth-network/'
@@
 
-4. In order to be able to mine, you need to create the DAG datasets as
+\item In order to be able to mine, you need to create the DAG datasets as
 follows (the creates files require approximately 2~GB disk space and
@@
 bash create_dag.sh
+\end{verbatim}  
+\end{enumerate}
+
+\section{Modifications}
+
+\subsection{Changes for limited hardware}
+\label{sec:limited-hardware}
+
+The target platform of the ARGoS-blockchain interface are robots with
+limited hardware. The mining algorithm Proof-of-Work requires a 1~GB
+dataset by default that must be stored in the memory. On a Pi-puck,
+Proof-of-Work does not run out-of-the-box due the large DAG file that
+has to be kept in memory. In order to reduce the DAG file size
+(warning: this will likely decrease the security of the network in a
+real-world deployment but is fine in a private testnet), modify the
+file \verb|go-ethereum/consensus/ethash/algorithm.go| as described below.
+
+1. Change
+\begin{verbatim}
+datasetInitBytes   = 1 << 30 // Bytes in dataset at genesis
+\end{verbatim}
+to
+\begin{verbatim}
+datasetInitBytes   = 1 << 20 // Bytes in dataset at genesis
+\end{verbatim}
+2. Comment out the following lines:
+\begin{verbatim}
+//if epoch < maxEpoch {
+	  //return cacheSizes[epoch]
+	  //}
+\end{verbatim}
+        and        
+\begin{verbatim}
+//if epoch < maxEpoch {
+	  //return datasetSizes[epoch]
+	  //}
 \end{verbatim}
@@
 The functions of the interface can be found in the file
-\texttt{/interface/generic_interface.cpp}.
+\verb|/interface/generic_interface.cpp|.
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{abbrvnat}
 \bibliography{library}

### 2020-09-04 article.tex
upd

@@
 
-The following setup steps describe the installation using Linux as
-operating system.
+The following setup steps describe the installation of the
+ARGoS-Blockchain interface using Linux as operating system.
 
@@
   plugin\footnote{\url{https://github.com/demiurge-project/argos3-epuck}}
-  (carefully read the instructions on the respective respositories)
+  (see installation instructions below)
 \end{itemize}
+
+\subsection{Installation of ARGoS and the ARGoS e-puck plugin}
+
+This section briefly describes the setup of ARGoS and the ARGoS e-puck
+plugin. Please carefully read the README files of the respective
+respositories for more information.
+
+1. Obtain the source code of both repositories
+
+\begin{verbatim}
+$ git clone https://github.com/ilpincy/argos3.git
+$ git clone https://github.com/demiurge-project/argos3-epuck.git
+\end{verbatim}
+
+2. Install the dependencies.
+
+\begin{verbatim}
+$ sudo apt-get install cmake libfreeimage-dev libfreeimageplus-dev \
+  qt5-default freeglut3-dev libxi-dev libxmu-dev liblua5.3-dev \
+  lua5.3 doxygen graphviz graphviz-dev asciidoc
+\end{verbatim}
+
+2. Build and install ARGoS system-wide.
+\begin{verbatim}
+$ cd argos3/
+$ mkdir build
+$ cd build
+$ cmake ../src
+$ make
+$ make doc
+$ sudo make install
+\end{verbatim}
+
+3. Build and install the ARGoS-epuck plugin.
+
+\begin{verbatim}
+$ cd ../../argos3-epuck/
+$ mkdir build
+$ cd build
+$ cmake ../src
+$ make
+$ sudo make install
+\end{verbatim}
 
@@
 \begin{verbatim}
-git clone https://github.com/Pold87/robot-swarms-need-blockchain"
+$ git clone https://github.com/Pold87/robot-swarms-need-blockchain
 \end{verbatim}
@@
 % New item
-\item Set the ARGoS folder in the file \verb|start_network.sh|.
+\item Set the ARGoS folder in the file \verb|global_config_blockchain.sh|.
 
 \begin{verbatim}
-$ nano ../local_scripts/start_network
+$ nano ../local_scripts/global_config_blockchain.sh
 ARGOSFOLDER="/home/vstrobel/software/interface/robot-swarms-need-blockchain"
 \end{verbatim}
-% TODO: Run script with NUMROBOTS=0
+
+\item In order to be able to mine, you need to create the DAG
+  datasets. Creating the files requires approximately 2~GB disk space
+  and the execution of the script can take take several minutes.
+
+\begin{verbatim}
+cd local_scripts/
+bash create_dag.sh
+\end{verbatim}
+
 \end{enumerate}
@@
 
+\begin{enumerate}
 \item Compile the ARGoS code.
@@
 \end{verbatim}
+\end{enumerate}
 
-\item In order to be able to mine, you need to create the DAG datasets as
-follows (the creates files require approximately 2~GB disk space and
-the execution of the script can take take several minutes):
+\section{Run}
+
+Usually, the network is created when a swarm robotics experiment is
+started, using one of the starter scripts in the folder
+\verb|robot-swarms-need-blockchain/starters/|. Start with a plain
+experiment to see if the installation was succesful. The command needs
+to be executed from the base folder of the ARGoS part of the
+blockchain interface.
 
 \begin{verbatim}
-cd local_scripts/
-bash create_dag.sh
+$ bash starters/1_Plain.sh 3
 \end{verbatim}
-\end{enumerate}
+
+You can also start the Ethereum network without ARGoS, using
+the following command:
+
+\begin{verbatim}
+$ bash local_scripts/start_network.sh <number of nodes>
+\end{verbatim}
+
+That is, \verb|bash local_scripts/start_network.sh 5|, would create a
+private Ethereum network with 5 nodes.
 
@@
 \end{verbatim}
-
-\section{Run}
-
-Usually, the network is created when a swarm robotics experiment is
-started, using one of the start scripts in
-\url{https://github.com/Pold87/robot-swarms-need-blockchain}.
-
-However, you can also start the Ethereum network without ARGoS, using
-the following command:
-
-\begin{verbatim}
-bash local_scripts/start_network.sh <number of nodes>
-\end{verbatim}
-
-That is, \verb|bash local_scripts/start_network.sh 5|, would create a
-private Ethereum network with 5 nodes.
 

### 2020-09-04 article.tex
fastpush

@@
 \title{ARGoS-Blockchain interface}
-\author{Volker Strobel \and Alex Pacheco \and Marco Dorigo}
+\author{Volker Strobel \and Eduardo Castell\'o Ferrer \and Alex Pacheco \and Marco Dorigo}
 
@@
 \begin{abstract}
-
   This technical report documents the installation and use of the
@@
 technology and swarm robotics
-\cite{StrCasDor2018:aamas,StrDor2018:ants,StrFerDor2017:techreport-013}. This
+\cite{PacStrDor2020:techreport-007,StrCasDor2018:aamas,StrDor2018:ants,StrFerDor2017:techreport-013}. This
 combination proves useful for security, consensus finding, robot
@@
 a suitable robot platform, such as the Pi-puck robot. Additionally, a
-suitable platform is important before running time- and cost-consuming
+suitable platform is important before running time-consuming
 experiments with real robots.

### 2020-09-11 article.tex
upd

--- 2020-09-04 article.tex
@@
 \title{ARGoS-Blockchain interface}
-\author{Volker Strobel \and Alex Pacheco \and Marco Dorigo}
+%\author{Volker Strobel \and Alex Pacheco \and Marco Dorigo}
 
@@
   ARGoS-blockchain interface.
-
 \end{abstract}
@@
 technology and swarm robotics
-\cite{StrCasDor2018:aamas,StrDor2018:ants,StrFerDor2017:techreport-013}. This
+\cite{StrCasDor2018:aamas,StrDor2018:ants,StrFerDor2017:techreport-013,Cas2016:arxiv,PacStrDor2020:techreport-001,PacStrDor2020:techreport-007}. This
 combination proves useful for security, consensus finding, robot
 economies, robot-as-a-service, and potentially many other applications
-in robot swarms. However, conducting experiments involving robot
+in robot swarms. Being a young research field, it misses standardized
+research platforms. However, conducting experiments involving robot
 swarms and blockchain technology requires a suitable testbed. A widely
@@
 experiments. Therefore, a simulation testbed is needed to quickly try
-and test different controllers and scenarios. The ARGoS robot
+and test different controllers and scenarios. The ARGoS robot swarm
 simulator \citep{PinTriOGr-etal2012:si} is the state-of-the-art
@@
 is available at
-\footnote{\url{https://github.com/Pold87/ARGoS-Blockchain-interface}}
+\footnote{\url{https://github.com/Pold87/AB-interface-Blockchain-module}}
 and the ARGoS module is available at
-\footnote{\url{https://github.com/Pold87/robot-swarms-need-blockchain}}.
+\footnote{\url{https://github.com/Pold87/AB-interface-ARGoS-module}}.
 
@@
 The following setup steps describe the installation of the
-ARGoS-Blockchain interface using Linux as operating system.
+ARGoS-Blockchain interface using Linux as operating system. There is a
+video tutorial at \url{https://www.youtube.com/watch?v=IqiTJK_uzUY}
+that described the entire installation process, starting from a blank
+Ubuntu 20.04 LTS installation.
 
@@
 \end{itemize}
+
+\subsection{Installation of Docker}
+
+The installation of the Docker Engine on Ubuntu is described at
+\url{https://docs.docker.com/engine/install/ubuntu/}. Below, you will
+find information about how to control the Docker Engine as a non-root
+user.
+
+\begin{enumerate}[leftmargin=*]
+\item By default, the execution of Docker commands requires root user
+  privileges. To facilitate the execution of Docker commands, it is
+  possible to run Docker as non-root user (i.e., without
+  \texttt{sudo}) as follows (see
+  \url{https://docs.docker.com/engine/install/linux-postinstall/} for
+  more detailed instructions).
+\begin{verbatim}
+$ sudo groupadd docker
+$ sudo usermod -aG docker $USER
+$ newgrp docker
+\end{verbatim}
+
+\item Additionally, restarting the Docker engine via \verb|systemctl|
+  requires sudo privileges. To control Docker as a normal user, add an
+  entry in the sudoers file.
+\begin{verbatim}
+$ sudo visudo -f /etc/sudoers.d/docker
+volker ALL= NOPASSWD: /bin/systemctl restart docker.service
+volker ALL= NOPASSWD: /bin/systemctl start docker.service
+volker ALL= NOPASSWD: /bin/systemctl stop docker.service
+\end{verbatim}
+\end{enumerate}
 
@@
 \begin{verbatim}
-$ sudo apt-get install cmake libfreeimage-dev libfreeimageplus-dev \
+$ sudo apt-get install cmake g++ libfreeimage-dev libfreeimageplus-dev \
   qt5-default freeglut3-dev libxi-dev libxmu-dev liblua5.3-dev \
@@
 $ sudo make install
+$ sudo ldconfig
 \end{verbatim}
 
-3. Build and install the ARGoS-epuck plugin.
+3. Verify that the installation was succesful.
+
+\begin{verbatim}
+$ argos3 --version
+ARGOS_VERSION=3.0.0-beta56
+ARGOS_INSTALL_PREFIX=/usr/local
+ARGOS_USE_DOUBLE=ON
+ARGOS_WITH_LUA=ON
+ARGOS_BUILD_FLAGS= -Wall -g -ggdb3
+\end{verbatim}
+
+4. Build and install the ARGoS-epuck plugin.
 
@@
 $ sudo make install
+$ sudo ldconfig
 \end{verbatim}
@@
 \begin{verbatim}
-$ mkdir interface/
-$ cd interface/
+$ mkdir AB-interface/
+$ cd AB-interface/
 $ pwd
-/home/vstrobel/software/interface
+/home/volker/software/AB-interface
 \end{verbatim}
@@
 \begin{verbatim}
-$ git clone https://github.com/Pold87/ARGoS-Blockchain-interface
+$ git clone https://github.com/Pold87/AB-interface-Blockchain-module.git \ 
+  Blockchain-module
 \end{verbatim}
@@
 \begin{verbatim}
-$ git clone https://github.com/Pold87/robot-swarms-need-blockchain
+$ git clone https://github.com/Pold87/AB-interface-ARGoS-module.git ARGoS-module
 \end{verbatim}
@@
 
-\begin{enumerate}[leftmargin=*]
-\item By default, the execution of Docker commands requires root user
-  privileges. To facilitate the execution of Docker commands, it is
-  possible to run Docker as non-root user (i.e., without
-  \texttt{sudo}) as follows (see
-  \url{https://docs.docker.com/engine/install/linux-postinstall/} for
-  more detailed instructions).
-\begin{verbatim}
-$ sudo groupadd docker
-$ sudo usermod -aG docker $USER
-\end{verbatim}
+\begin{enumerate}
  % New item
 \item Obtain the Ethereum source code and store it in the folder
-\verb|myowngeth/|. Please note that Ethereum is compiled locally
+\verb|custom-geth/|. Please note that Ethereum is compiled locally
 here. While there are Docker images for Ethereum, this allows for more
@@
 $ cd ARGoS-Blockchain-interface/ && cd geth/
-$ git clone https://github.com/ethereum/go-ethereum.git myowngeth/
+$ git clone https://github.com/ethereum/go-ethereum.git custom-geth/
 \end{verbatim}
@@
 % New item
-\item Set the ARGoS folder in the file \verb|global_config_blockchain.sh|.
+\item Next, we need to link the blockchain module and the ARGoS module. First, set the ARGoS folder in the file \verb|global_config_blockchain.sh|.
 
@@
 $ nano ../local_scripts/global_config_blockchain.sh
-ARGOSFOLDER="/home/vstrobel/software/interface/robot-swarms-need-blockchain"
+ARGOSFOLDER="/home/volker/software/AB-interface/ARGoS-module"
+\end{verbatim}
+
+\item Then, set the variable \verb|DOCKERBASE| in the file
+  \verb|global_config.sh| of the ARGoS module (Module~2) to the full
+  path where the blockchain module (Module~1) repository is located on
+  your computer, for example:
+
+\begin{verbatim}
+$ nano /home/volker/software/AB-interface/ARGoS-module/global_config.sh
+DOCKERBASE='/home/volker/software/AB-interface/Blockchain-module/'
 \end{verbatim}
@@
   datasets. Creating the files requires approximately 2~GB disk space
-  and the execution of the script can take take several minutes.
+  and the execution of the script can take take several minutes. If
+  you want to run the software on limited hardware, you can decrease
+  the size of the DAG dataset by carrying out the modifcations
+  described in Section~\ref{sec:limited-hardware}.
 
@@
 \begin{verbatim}
-$ cd /home/vstrobel/software/interface/robot-swarms-need-blockchain/
+$ cd /home/volker/software/AB-interface/ARGoS-module/
 $ mkdir build/
@@
 $ make
-\end{verbatim}
-
-\item Set the variable \verb|DOCKERBASE| in the file
-\verb|global_config.sh| to the full path where the Blockchain module
-(Module 1) repository is located on your computer, for example:
-
-\begin{verbatim}
-$ cd ..
-$ nano global_config.sh
-DOCKERBASE='/home/vstrobel/Documents/docker-geth-network/'
 \end{verbatim}
@@
 started, using one of the starter scripts in the folder
-\verb|robot-swarms-need-blockchain/starters/|. Start with a plain
+\verb|ARGoS-module/starters/|. Start with a plain
 experiment to see if the installation was succesful. The command needs
@@
 \begin{verbatim}
-$ bash starters/1_Plain.sh 3
+$ bash starters/1_Plain.sh
 \end{verbatim}
@@
 \section{Modifications}
+\label{sec:modifications}
 
@@
 individual containers for the robots are running:
+%
+\begin{verbatim}
+$ docker service ps ethereum_bootstrap --no-trunc
+$ docker service ps ethereum_eth --no-trunc
+\end{verbatim}
+%
+
+% \begin{verbatim}
+% ID                          NAME                   IMAGE               NODE                DESIRED STATE       CURRENT STATE           ERROR               PORTS
+% ik6isqahrmsgzkzkj4pbb2zd5   ethereum_bootstrap.1   mygeth:latest       volker-TERRA-PC     Running             Running 4 minutes ago   
+% \end{verbatim}
+
+If the value of \verb|DESIRED STATE| is \verb|RUNNING|, take a look at
+the log files of the containers.
 
 \begin{verbatim}
-
+$ docker logs ethereum_bootstrap.1.ik6isqahrmsgzkzkj4pbb2zd5
 \end{verbatim}
-
-% Mention --no-trunc and how to access logs
 
@@
 \item \texttt{removePeer}: Add a peer based on its enode.
-\item \texttt{scInterface}: 
+\item \texttt{scInterface}: A generic smart contract interface.
+\item \texttt{scReturn0(string function, long long wei)}: 
 \end{itemize}
+
+\subsection{Compilation}
+\label{sec:compilation}
+
+After changes to the ARGoS files are made, the code needs to be
+recompiled.
+
+\begin{verbatim}
+$ cd build/
+$ make
+\end{verbatim}
 

### 2020-09-12 article.tex
corrected folder

@@
 \title{ARGoS-Blockchain interface}
-%\author{Volker Strobel \and Alex Pacheco \and Marco Dorigo}
+%\author{Volker Strobel \and Eduardo Castello Ferrer \ and Alex Pacheco \and Marco Dorigo}
 
@@
 \begin{verbatim}
-$ cd ARGoS-Blockchain-interface/ && cd geth/
+$ cd Blockchain-module/ && cd geth/
 $ git clone https://github.com/ethereum/go-ethereum.git custom-geth/

### 2020-09-12 article.tex
code easier to copy

@@
   requires sudo privileges. To control Docker as a normal user, add an
-  entry in the sudoers file.
+  entry in the sudoers file (replace \verb|volker| with your user
+  name).
 \begin{verbatim}
 $ sudo visudo -f /etc/sudoers.d/docker
+\end{verbatim}
+\begin{verbatim}
 volker ALL= NOPASSWD: /bin/systemctl restart docker.service
